---
title: "Lab7 Assignment"
author: "Gabriela Garza-Long"
date: "04-20-2025"
output: html_document
---

``````{r}
library(data.table)
library(haven)
library(foreign)
library(ggplot2)
library(dplyr)

#1. Load Household Dataset
hts <- read_sav("~/Desktop/Methods2_Labs/07-lab-assignment-2025-Green-range/datasets/HTS.household.10regions.sav")
hts <- as.data.table(hts)

# Convert and clean
hts$hhincome <- as.numeric(hts$hhincome)
hts_clean <- hts[!is.na(hhincome)]
hts_clean$htype <- as.factor(hts_clean$htype)

# Filter only two groups
hts_two_groups <- hts_clean[htype %in% c("0", "1")]

# Run t-test
t.test(hhincome ~ htype, data = hts_two_groups)

``````

#2. T-Test: Income by Housing Type

``````{r}
# Convert and clean
hts$hhincome <- as.numeric(hts$hhincome)
hts_clean <- hts[!is.na(hhincome)]
hts_clean$htype <- as.factor(hts_clean$htype)

# Filter only two groups
hts_two_groups <- hts_clean[htype %in% c("0", "1")]

# Run t-test
t.test(hhincome ~ htype, data = hts_two_groups)

#2. T-Test: Income by Housing Type

# Convert and clean
hts$hhincome <- as.numeric(hts$hhincome)
hts_clean <- hts[!is.na(hhincome)]
hts_clean$htype <- as.factor(hts_clean$htype)

# Filter only two groups
hts_two_groups <- hts_clean[htype %in% c("0", "1")]

# Run t-test
t.test(hhincome ~ htype, data = hts_two_groups)

#Histogram of Income by Housing Type

ggplot(hts_two_groups, aes(x = hhincome, fill = htype)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 30) +
  labs(title = "Histogram of Household Income by Housing Type",
       x = "Household Income", fill = "Housing Type") +
  theme_minimal()

``````

#3. T-Test: San Antonio 
``````{r}
library(data.table)
library(haven)
library(foreign)
library(ggplot2)
library(dplyr)

#1. Load Household Dataset
hts <- read_sav("~/Desktop/Methods2_Labs/07-lab-assignment-2025-Green-range/datasets/HTS.household.10regions.sav")
hts <- as.data.table(hts)

# Convert and clean
hts$hhincome <- as.numeric(hts$hhincome)
hts_clean <- hts[!is.na(hhincome)]
hts_clean$htype <- as.factor(hts_clean$htype)

# Filter only two groups
hts_two_groups <- hts_clean[htype %in% c("0", "1")]

# Run t-test
t.test(hhincome ~ htype, data = hts_two_groups)

``````
#2. T-Test: Income by Housing Type

``````{r}
library(data.table)
library(haven)
library(foreign)
library(ggplot2)
library(dplyr)

#1. Load Household Dataset
hts <- read_sav("~/Desktop/Methods2_Labs/07-lab-assignment-2025-Green-range/datasets/HTS.household.10regions.sav")
hts <- as.data.table(hts)

# Convert and clean
hts$hhincome <- as.numeric(hts$hhincome)
hts_clean <- hts[!is.na(hhincome)]
hts_clean$htype <- as.factor(hts_clean$htype)

# Filter only two groups
hts_two_groups <- hts_clean[htype %in% c("0", "1")]

# Run t-test
t.test(hhincome ~ htype, data = hts_two_groups)

``````

#3. T-Test: San Antonio 
``````{r}
# Filter San Antonio
hts_SA <- hts %>% filter(region == 15)

# Create jobpop group
jobpop_median <- median(hts_SA$jobpop, na.rm = TRUE)
hts_SA <- hts_SA %>%
  mutate(jobpop_group = ifelse(jobpop > jobpop_median, "High", "Low")) %>%
  filter(!is.na(lnvmt), !is.na(jobpop_group))

# Run t-test
t.test(lnvmt ~ jobpop_group, data = hts_SA)

``````


#4. ANOVA: lnvmt by Income Category
``````{r}
# Ensure categorical income
hts_SA$income_cat <- as.factor(hts_SA$income_cat)
anova_result <- aov(lnvmt ~ income_cat, data = hts_SA)
summary(anova_result)

#QQ Plot (Normality)
qqnorm(hts_SA$lnvmt)
qqline(hts_SA$lnvmt)

#Boxplot of lnvmt by Income

boxplot(lnvmt ~ income_cat, data = hts_SA,
        main = "Boxplot of Log Vehicle Miles Traveled",
        xlab = "Income Category", ylab = "lnvmt")

``````


#5. Tukey Post Hoc Test
``````{r}
tukey_result <- TukeyHSD(anova_result)
plot(tukey_result)
tukey_result

``````


#6. Simulated T-Test Plot
``````{r}
# Create single_family var
hts[, single_family := ifelse(htype == 0, "Single Family", "Other")]
hts[, single_family := factor(single_family)]
t_test_income <- t.test(hhincome ~ single_family, data = hts)

# Plot
curve(dt(x, df = t_test_income$parameter),
      from = -4, to = 4, col = "darkblue", lwd = 2,
      main = "Simulated T-distribution (T-Test)",
      xlab = "T-value", ylab = "Density")
abline(v = t_test_income$statistic, col = "red", lwd = 2)
text(t_test_income$statistic, 0.05,
     labels = paste0("t = ", round(t_test_income$statistic, 2)),
     pos = 4, col = "red")
``````


#7. Summary of Results
#T-test 1: There is a statistically significant difference in household income between housing types (t = ..., p < 0.001).

#T-test 2: Households in high job-population areas drive less on average than those in low areas (p < 0.05).

#ANOVA: Significant variation in VMT by income category (p < 0.01); Tukey’s test confirms differences between low and high income groups.
